---
title: "writing_functions"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)
library(readxl)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
source("sim_mean_sd.R")
```

we can simulate by running our function
```{r}
sim_mean_sd(n_subj = 400)

```

can I verify the central limit theorem?

First, do it with a for loop
for 100 trials, simulate mean and sd
```{r}
output = vector("list", length = 100)

for (i in 1:100) {
  
  output[[i]] = sim_mean_sd(30)
  
}

output |> 
  bind_rows() |> 
  ggplot(aes(x = mu_hat)) + 
  geom_density()
```


map statment
```{r}
sim_results_df = 
  expand_grid (
    sample_size = 30,
    iter = 1:100
    )|> 
  mutate(
    results = map(sample_size, sim_mean_sd)
  ) |> 
  unnest(results)  
```

expand grid - every possible combination of iteration 1-100 and same size

```{r}
sim_results_df = 
  expand_grid (
    sample_size = c(30,60,90,120),
    iter = 1:100
    )|> 
  mutate(
    results = map(sample_size, sim_mean_sd, n_subj = 30)
  ) |> 
  unnest(results)
```



Let's try to summarize ..

```{r}
sim_results_df |> 
  group_by(sample_size) |> 
  summarize(
    emp_mean = mean(mu_hat),
    emp_se = sd(mu_hat)
  )
```

## Simple Linear Regression

this is how you fit linear regression
```{r}
sim_df = 
  tibble(
    x = rnorm(30, mean = 1, sd = 1),
    y = 2 + 3 *x +rnorm(30, 0, 1)
  )

sim_df |> 
  ggplot(aes(x=x, y=y))+
  geom_point()

slr_fit = lm(y~x, data = sim_df)

coef(slr_fit)
```

we want to fit linear regression over and over again.
turn the above into function

```{r}
sim_regression = function (n_subj, beta_0 = 2, beta_1 = 3) {
  
  sim_df = 
  tibble(
    x = rnorm(n_subj, mean = 1, sd = 1),
    y = beta_0 + beta_1 * x + rnorm(n_subj, 0, 1)
  )


  slr_fit = lm(y~x, data = sim_df)

  tibble(
    beta0_hat = coef(slr_fit)[1],
    beta1_hat = coef(slr_fit)[2]
  )

}
```


```{r}
sim_regression(n_subj = 30)
```

```{r}
vector("list", length = 500)

for(i in 1:500) {
  output[[i]] = sim_regression(n_subj = 30)
}

output |> 
  bind_rows()
```

column that keeps track of iteration number
```{r}
slr_sim_results_df = 
  expand_grid(
    sample_size = 30,
    iter = 1:500
  ) |> 
  mutate(
    results = map(sample_size, sim_regression)
  ) |> 
  unnest(results)

slr_sim_results_df |> 
  ggplot(aes(x = beta0_hat, y = beta1_hat)) +
  geom_point()
```

  # when intercept is higher, slp is lower and vice versa
  
  
## one more example

birthday problem - how many people do you need to put in a room to have 50% chance that 2 people have the same birthday

```{r}
birthdays = sample(1:365, 5, replace = TRUE) # you can pick same number multiple times

unique(birthdays) < 5
unique(c(1,2,3,4,5)) # the unique values that exist here
length(unique(birthdays)) < 5 # if repeat this statement is T if no repeat then statement is false

repeated_bday = length(unique(birthdays)) < 5
```

putting above into a function

```{r}
bday_sim = function(n_room) {
  
  birthdays = sample(1:365, n_room, replace = TRUE)
  
  repeated_bday = length(unique(birthdays)) < n_room
  
  repeated_bday 
}

bday_sim(20)
```

want to repeat the function above multiple times

```{r}
bday_sim_results = 
  expand_grid(
    bdays = 5:50,
    iter = 1:50
  ) |> 
  mutate(
    result = map_lgl(bdays, bday_sim)
  ) |> 
  group_by(
    bdays
  ) |> 
  summarize(
    prob_repeat = mean(result)
  )
```

```{r}
bday_sim_results |> 
  ggplot(aes(x = bdays, y = prob_repeat)) +
  geom_point() +
  geom_line()
```

if you only see sig results, your study is underpowered
